Training a Gaming Agent on Brainwaves

### 2020-10-15 review2/response.letter.tex

\vskip+1ex
\noindent \dotfill

\section*{\fbox{AE Transcript:}}

AE's Comments to Author:

Associate Editor
Comments to the Author:

The paper is improved but would benefit from a further update to

\begin{itemize}
\item improve abstract
\item provide further interpretation/discussion of findings
\item fix minor errors/typos
\item address the grand average recommendation
\end{itemize}

@@

\section*{\fbox{Reviewer \#2 Transcript:}}

Comments to the Author
The authors have satisfactorily addressed my concerns. I have a minor comment regarding Figure 3. It is unclear what the brown bar labelled GrandAvg represents since there is one for each participant. Please clarify. I suggest having another set of bars (a ninth) for the grand average (average over participants) for each of the algorithms used.

\subsection*{\ovalbox{AE General Comments}}
AE's Comments to Author:

Associate Editor
Comments to the Author:
The paper is improved but would benefit from a further update to

- improve abstract

\begin{quotation}
{\color{blue}
We have rewrote the abstract explaining in further details the obtained results.  
}
\end{quotation}

@@

- provide further interpretation/discussion of findings

\begin{quotation}
{\color{blue}
We included extra information and verified the contributions/findings.  This information was included in the abstract as well as in the Conclusion (Section IV).
}
\end{quotation}

@@

- fix minor errors/typos

\begin{quotation}
{\color{blue}
We fixed the carefully reported issues gently provided by Reviewers.
}
\end{quotation}

- address the grand average recommendation

\begin{quotation}
{\color{blue}
We addressed this issue.  We think there could be a misunderstanding from our side about the suggestion/recommendation of the Reviewer.  We added additional information in this response letter, and apologize for this.
}
\end{quotation}

@@

Line 40 “This information is used to make a gaming agent $<$that$>$ improves ....”

\begin{quotation}
{\color{blue}
This text is no longer contained in the manuscript.  We verified throughout the Manuscript to avoid similar mistakes.
}
\end{quotation}

Line 11 col 2 “how biological agents learn from its $<$their$>$ environment by exploring it and getting feedback rewards, either negative or positive.”

\begin{quotation}
{\color{blue}
We fixed this grammatical error.  Thank you very much for noticing it.
}
\end{quotation}

@@

Page 4
Lin 24 This allows \sout{to learn} the Q-Table $<$to learn$>$ based on the subject’s feedback from the movements the agent
took, which are chosen pseudo-randomly, while executing the brainwave session.

\begin{quotation}
{\color{blue}
Fixed issue.  We appreciate a lot the level of detail in the comments.
}
\end{quotation}

Line 50 “The best overall performance is obtained using Logistic Regression.” // also need to discuss the significance of the overall levels of accuracy in Discussion. Is a best performance of 0.672 useful/acceptable? How does it compare to other researchers (if a comparison is possible).

\begin{quotation}
{\color{blue}
On Section IV, third paragraph we tackled this issue.  The overall performance of 0.672 is low.  Although a straightforward comparison could be misleading in the context of the ErrP experiment, we added a brief recount of the values obtained for other similar works.  We also emphasized this point in future works at the end of the Conclusions section.
}
\end{quotation}

@@

Line 42 col 2 “These results are also consistent with their classification ROC curves, shown \sout{on} $<$in$>$ Figures 6 obtained for both subjects, where the area under the curve are close to chance level.” // Why didn’t you show one good ROC curve and one bad one. This would provide the basis for comparison.

\begin{quotation}
{\color{blue}
Due to previous Reviewer's requests, we added all the ROC curves on the new Figure 6.  We now emphasized on this figure's caption which ROC curves where showing an effective identification of the ErrP potential and which did not.  We added this in the text highlighting the important finding that we could not provide an enhancement of the agent performance based on the rewards from subjects where the obtained ROC curves were not good (Section III, page 5).
}
\end{quotation}

@@

Page 5
Figure 5 “Y axis shows the averaged number of steps, while x axis show the number of
experiences used to cumulative train the Q-Table.”
// Why does Fig5 E only have 2 sets of data? Some discussion of the (smaller) change from 1 – 2 would be appropriate (e.g. I Discussion)
Line 51 “No\sout{t} performance gain is evidenced, the agents learn nothing which implies that the reward information is useless.” // useless is not a good choice of wording –provides no value

\begin{quotation}
{\color{blue}
We fixed the inappropriate wording, and appreciate the Reviewer for pointing this out.  
Figure 5, for Subject 5 (in the first version of the manuscript it was labeled E) has only two sets of data because that Participant (OHC) couldn't complete more sessions.  Hence we used two training sessions to calculate the Q-Table and subsequently performed the run session (i.e. the agent performs movements randomly based on the trained QTable for 200 iterations).
}
\end{quotation}

Line 55 “It can be seen that the overall performance of the agent improves as long as there are more experiences to be used to train it, regardless if they were generated from the brainwaves classification from different subjects.” // more explanation/discussion needed. Is this an average effect from positive learners?

\begin{quotation}
{\color{blue}
This is a great question.  Traditional transfer learning strategies focus on the procedure to enrich a classifier training it with a vast dataset and using it to generalize to new data from unseen scenarios.  While dealing with EEG data, this strategy tends to fail, and other approaches are required.  We found in this work, that even when that strategy fails, if the system is coupled with a basic reinforcement learning algorithm, the cumulative rewards from different subjects, can be used to effectively transmit information and improve the agent performance.  We agree with the Reviewer that this may be due to the average effect of positive or learners with a higher accuracy, but we think it may be related to the fact that the RL algorithm is quite robust and if we can get a higher specificity it will lead to eventually learn, and the cumulative usage of data from various OHCs help to overcome this issue and improve the agent performance.
}
\end{quotation}

@@

This work aims to state whether ErrP signals could be used to train a gaming agent using reinforcement learning. The collected data show that ErrP signals can in fact be classified and used to train an agent effectively. // Can you link these findings to confirm/challenge other recent research in Err potentials

\begin{quotation}
{\color{blue}

These results are supported by the literature. As we mention in the Introduction, similar experiments have been performed (R. Chavarriaga, A. Sobolewski, and J. d. R. Millán, "Errare machinale est: The use of error-related potentials in brain-machine interfaces," Frontiers in Neuroscience,2014).  Our novel approach is to use the rewards to train a gaming agent. 
We have added this confirmation in the Conclusion Section IV.
}
\end{quotation}

@@

Page 6
Fig 8 – why use data from subject 6 (a non-learner)?

\begin{quotation}
{\color{blue}
The experiment that we performed to generate the Figure 9 doesn't include data from Subject 5 and 6. This is emphasized on the Figure's caption and on the main text in the Result Section.  In the Figure 9, the X axis show the progressive number of gaming agent training matches used to update the Q-Table (in this case they are from different subjects).
}
\end{quotation}

@@

Line 42 “However, even though this implies that the agent misses frequently that an action taken is wrong, this is not hindering the overall performance and the agent is still learning” //

Line 20 col 2 Despite that, the rewards generated from different subjects can be used to train the same Q- Table to improve its performance, which may lead to strategies where the overall performance is improved based on the information from different human critics at the same time.
//These are key findings, worth more discussion/interpretation and inclusion in abstract.

\begin{quotation}
{\color{blue}
We included the both finding in the abstract and we emphasized them  in Section IV. 
}
\end{quotation}

### 2020-10-15 ErrP.tex
Le doy una vuelta más de rosca. Ya casi está.

% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\begin{abstract}
%objective
Error-related potential (ErrP) are a particular type of ERP elicited by a person attending a recognizable error.  These signals can be used to train a gaming agent by a Reinforcement Learning (RL) algorithm to learn an optimal policy.  The experimental process consists of an observational human critic (OHC) observing a simple game scenario while their Electroencephalographic (EEG) brain signals are captured. The game consists of a grid, where a green spot has to reach a desired target in the fewest amount of steps. 
%Results
% Expand the results of the paper as this is the contribution. “Results show that there is an effective transfer of information and that the agent learns successfully to solve the game eﬀiciently. Both the underlined terms need to cite evidence from the paper.
% The rewards generated from different subjects can be used to train the same Q- Table to improve its performance, which may lead to strategies where the overall performance is improved based on the information from different human critics at the same time. //These are key findings, worth more discussion/interpretation and inclusion in abstract.
Results show that there is an effective transfer of information and that the agent successfully learns to solve the game efficiently, from the initial 97 steps on average required to reach the target to the optimal number of 8 steps.  Our results are expressed in threefold: (i) the structure of a simple grid-based game that can elicit the ErrP signal component, (ii) the verification that the reward function only penalizes wrong steps, which means that type II error (not properly identifying a wrong movement) does not affect significantly the agent learning process; (iii) collaborative rewards from multiple observational human critics can be used to train the algorithm effectively and can compensate the lack of low accuracy (approximately 67\%) or the limited scope of transfer learning schemes.

%The algorithm can be trained with information from different OHCs, since the resulting dataset is composed of rewards for each corresponding step, independent of who they were generated from. This shows that the error classification accuracy (approximately 0.67) is good enough for the algorithm to learn. The reward function only penalizes wrong steps, which means that type II error (not properly identifying a wrong movement) does not affect the accuracy, they only make the learning process slower.
%conclusions
%Our results are expressed in threefold (i) the structure of a simple grid-based game that can elicit the ErrP signal component; (ii) the verification that low classification accuracy of just above chance level that produces noisy rewards is enough to allow an agent to learn the optimal policy; (iii) collaborative rewards from multiple observational human critics can compensate the lack of accuracy or the limited scope of transfer learning schemes.

@@

Therefore,  we aim to use the information extracted from brainwaves to enhance the performance of a gaming agent.  The three contributions are (1) a simple game mechanics and agent that can elicit the ErrP potential, (2) results that confirm that even when ErrP classification accuracy is low but with a high specificity and produces a noisy reward signal, enough information is generated for an agent to learn the optimal policy and solve a simple game and (3) collaborative rewards from multiple observational human observers can compensate the lack of classification accuracy or the inefficacy of transfer learning procedures for brainwaves signals.

@@

In Section \ref{section:materials} the general layout of the cognitive game is described. Sections \ref{brainwavesession} and \ref{cognitive_experiment_details} outline the cognitive game procedure used to obtain rewards in the form of ErrP components. Section \ref{q_learning_step_alg} describes the gaming agent learning procedure.  Lastly, results and conclusions are exposed in Sections \ref{results} and \ref{conclusions}.

@@

This phase is divided into a sequence of run sessions and a gaming agent training match.  During the run session, the agent plays 200 matches guided by a specific Q-Table with a 5\% chance of randomly selecting a movement, to reduce deadlocks and loops.  Following the run session, the agent performs a single gaming agent training match.  The gaming agent starts first with a Q-Table initialized with zeros, so the initial policy for the agent is randomized.  For the agent to learn from the feedback generated by the OHC, movement actions are determined by the reply of the agent's actions that were taken during one brainwave session match, in an offline reinforcement learning scheme~\cite{Levine2020}. This allows the Q-Table to be built based on the OHC's feedback from the movements the agent took, which were executed pseudo-randomly during the brainwave session.  The previously mentioned feedback is not explicit as it comes from the interpreted brain signal data. This implies that the reward is determined by the OHC's brain activity.

@@

\begin{figure}[ht]
    \centering
    \includegraphics[width=0.5\textwidth]{revisedimages/avg_closer.eps}
    \includegraphics[width=0.5\textwidth]{revisedimages/avg_further.eps}
        \includegraphics[scale=1.0]{revisedimages/electrodes.png}
    \caption{Grand Average of 2-seconds segments for all OHCs for the "move closer" (top) and "move further" (middle) condition.  The increased height of the first peak around 0.4 seconds reflects the Error Potential response particularly on Pz, based on the electrode layout (bottom).}
    \label{fig:classifiers}
\end{figure}

Figure~\ref{fig:classifiers} shows the binary classification accuracy obtained for the eight OHCs using five different classification algorithms and using a 10-fold cross-validation procedure.  The best overall performance is obtained using Logistic Regression.

@@

On the other hand, Figure \ref{fig:avg_steps} shows the average amount of steps it takes for the agent to reach the goal for each OHC, as the Q-Table is progressively trained using the reward information obtained from the prediction of the trained classifier. Each point corresponds to a run session where the average number of steps for the agent to reach the goal using a specific Q-Table is specified, for 200 repetitions. The first point, at the x-value 0, represents the number of steps the agent takes to reach the goal with an untrained Q-Table, where movements are decided randomly. The next point corresponds to the amount of steps it takes to reach the goal using a policy derived from a Q-Table trained after one brainwave session match, and so on.

@@

Each time the gaming agent plays this simple game, it takes on average around 100 steps to reach the target spot.  We trained a classifier to recognize Error Potential from observational human critics that watch the agent playing the game.  We let the agent play again and we mark movements that trigger an error potential from the human critic.  We used those movements as rewards in a Reinforcement Learning scheme, and use them to train a Q-Table.  When we let the agent plays the game again, the number of steps that requires to solve the game is now reduced.  If we provide feedback based on random signals, we verified that no reduction is achieved and the average number of steps does not change.   This shows that there is an effective transfer of information from the brainwaves to the agent.   As this process is repeated, the agent keeps improving solving the game effectively, i.e. performing the minimum number of required steps to reach the goal.

@@

This work aims to propose a simple game mechanics that can use the ErrP component to train a gaming agent using a RL model. The collected data and the shows that ErrP signals can in fact be classified and used to train an agent effectively.

This proposal tries to keep the system as simple as possible, emphasizing information flow from the subjective error perception of the human critic. Rewards are generated using the signal processing and classification pipeline, and the Q-Table updates, enhancing the performance of the gaming agent.

One additional aspect to remark is the robustness of the learning strategy based on Q-Learning~\cite{Bauer2015,Rubin2012}.   The obtained accuracy to discriminate ErrPs is low, though on the same level to other similar results~\cite{Iturrate2013,Ehrlich2016}.  In this regard, dispite of the subjectivity of the ErrP response in terms of the proposed experiment, levels of 90\% of accuracy are reported~\cite{Chavarriaga2014}.  However, even with such low accuracy values, the RL algorithm was able to extract meaningful information from rewards that were helpful to improve, and often maximize, the agent's performance.  Additionally, one important aspect of the classification results is the low percentage of false positives (Figure~\ref{fig:confusionmatrix}), showing a high specificity. On the other hand, the percentage of false negatives is generally higher.  However, even though this implies that the agent misses frequently when a wrong action takes place, this is not hindering the overall performance and the agent is still learning. Though scarce, accurate rewards are very useful for the RL algorithm.
%However, even though this imply that the agent misses frequently that an action taken is wrong, this do not seems to be a serious issue because the agent improves its performance.

>>>>

% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\begin{abstract}
%objective
Error-related potential (ErrP) are a particular type of ERP elicited by a person attending a recognizable error.  These Electroencephalographic (EEG)  signals can be used to train a gaming agent by a Reinforcement Learning (RL) algorithm to learn an optimal policy.  The experimental process consists of an observational human critic (OHC) observing a simple game scenario while their brain signals are captured. The game consists of a grid, where a green spot has to reach a desired target in the fewest amount of steps.  Results show that there is an effective transfer of information and that the agent successfully learns to solve the game efficiently, from the initial 97 steps on average required to reach the target to the optimal number of 8 steps.  Our results are expressed in threefold: (i) the mechanics of a simple grid-based game that can elicit the ErrP signal component, (ii) the verification that the reward function only penalizes wrong steps, which means that type II error (not properly identifying a wrong movement) does not affect significantly the agent learning process; (iii) collaborative rewards from multiple observational human critics can be used to train the algorithm effectively and can compensate low classification accuracies (around 67\%) or the limited scope of transfer learning schemes.
%Results
% Expand the results of the paper as this is the contribution. “Results show that there is an effective transfer of information and that the agent learns successfully to solve the game eﬀiciently. Both the underlined terms need to cite evidence from the paper.
% The rewards generated from different subjects can be used to train the same Q- Table to improve its performance, which may lead to strategies where the overall performance is improved based on the information from different human critics at the same time. //These are key findings, worth more discussion/interpretation and inclusion in abstract.
%The algorithm can be trained with information from different OHCs, since the resulting dataset is composed of rewards for each corresponding step, independent of who they were generated from. This shows that the error classification accuracy (approximately 0.67) is good enough for the algorithm to learn. The reward function only penalizes wrong steps, which means that type II error (not properly identifying a wrong movement) does not affect the accuracy, they only make the learning process slower.
%conclusions
%Our results are expressed in threefold (i) the structure of a simple grid-based game that can elicit the ErrP signal component; (ii) the verification that low classification accuracy of just above chance level that produces noisy rewards is enough to allow an agent to learn the optimal policy; (iii) collaborative rewards from multiple observational human critics can compensate the lack of accuracy or the limited scope of transfer learning schemes.

@@

Therefore,  we aim to use the information extracted from brainwaves to enhance the performance of a gaming agent.  The three contributions are (1) a simple game mechanics and agent that can elicit the ErrP potential, (2) results that confirm that even when ErrP classification accuracy is low but with a high specificity, enough information is generated for an agent to learn the optimal policy and solve a simple game and (3) collaborative rewards from multiple observational human observers can compensate the lack of classification accuracy or the inefficacy of transfer learning procedures for brainwaves signals.

@@

This work unfold as follows: in Section \ref{section:materials} the general layout of the cognitive game is described. Sections \ref{brainwavesession} and \ref{cognitive_experiment_details} outline the cognitive game procedure used to obtain rewards in the form of ErrP components. Section \ref{q_learning_step_alg} describes the gaming agent learning procedure.  Lastly, results and conclusions are exposed in Sections \ref{results} and \ref{conclusions}.

@@

This phase is divided into a sequence of run sessions and gaming agent training matches.  During the run session, the agent plays 200 matches guided by a specific Q-Table with a 5\% chance of randomly selecting a movement, to reduce deadlocks and loops.  Following the run session, the agent performs a single gaming agent training match.  The gaming agent starts first with a Q-Table initialized with zeros, so the initial policy for the agent is randomized.  For the agent to learn from the feedback generated by the OHC, movement actions are determined by the reply of the agent's actions that were taken during one brainwave session match, in an offline reinforcement learning scheme~\cite{Levine2020}. This allows the Q-Table to be learned based on the OHC's feedback from the movements the agent took, which were executed pseudo-randomly during the brainwave session.  The previously mentioned feedback is not explicit as it comes from the interpreted brain signal data. This implies that the reward is determined by the OHC's brain activity.

@@

Grand averaged signal segments of 2-seconds length for all the OHCs can be seen on Figure~\ref{fig:grandaverage}.  The ErrP can be noticed more clearly on Pareto-central areas (Pz electrode), around 0.4~s with a more prominent positive peak.

\begin{figure}[ht]
    \centering
    \includegraphics[width=0.5\textwidth]{revisedimages/avg_closer.eps}
    \includegraphics[width=0.5\textwidth]{revisedimages/avg_further.eps}
        \includegraphics[scale=1.0]{revisedimages/electrodes.png}
    \caption{Grand average of 2-seconds segments for all OHCs for the "move closer" (top) and "move further" (middle) condition.  The increased height of the first peak around 0.4 seconds reflects the Error Potential response particularly on Pz, based on the electrode layout (bottom).}
    \label{fig:grandaverage}
\end{figure}

Figure~\ref{fig:classifiers}, on the other hand, shows the binary classification accuracy obtained for the eight OHCs using five different classification algorithms and using a 10-fold cross-validation procedure.  The best overall performance is obtained using Logistic Regression.

@@

Complementary, Figure \ref{fig:avg_steps} shows the average amount of steps it takes for the agent to reach the goal for each OHC, as the Q-Table is progressively trained using the reward information obtained from the prediction of the trained classifier. Each point corresponds to a run session where the average number of steps for the agent to reach the goal using a specific Q-Table is specified, for 200 repetitions. The first point, at the x-value 0, represents the number of steps the agent takes to reach the goal with an untrained Q-Table, where movements are decided randomly. The next point corresponds to the amount of steps it takes to reach the goal using a policy derived from a Q-Table trained after one brainwave session match, and so on.

@@

Each time the gaming agent plays this simple game, it takes on average around 100 steps to reach the target spot.  We trained a classifier to recognize Error Potential from observational human critics that observe the agent playing the game.  The gaming agent plays randomly and movements are marked triggered by the identification of an error potential from the human critic.  Those movements are used as rewards in a Reinforcement Learning scheme to train a Q-Table.  The gaming agent plays the game again, but this time the number of required steps to solve the game is reduced.  If rewards are provided based on random signals,  no reduction is achieved and the average number of steps does not change.   This shows that there is an effective transfer of information from the brainwaves to the agent.   As this process is repeated, the agent keeps improving solving the game effectively, i.e. performing the minimum number of required steps to reach the goal.

@@

This work aims to propose a simple game mechanics that can use the ErrP component to train a gaming agent using a RL model. As described in the literature~\cite{Chavarriaga2014}, ErrPs can be used to transmit subjective feedback to a computer.  The collected data shows that ErrP signals can in fact be classified and used to train an agent effectively.  This proposal tries to keep the system as simple as possible, emphasizing information flow from the subjective error perception of the human critic. Rewards are generated using the signal processing and classification pipeline, and the Q-Table updates, enhancing the performance of the gaming agent.

One additional aspect to remark is the robustness of the learning strategy based on Q-Learning~\cite{Bauer2015,Rubin2012}.   The obtained accuracy to discriminate ErrPs is low, though on the same level to other similar results~\cite{Iturrate2013,Ehrlich2016}.  In this regard, despite the subjectivity of the ErrP response in terms of the proposed experiment, levels of 90\%  accuracy are reported~\cite{Chavarriaga2014} as well.  Despite the low accuracy  in ErrP identification, the RL algorithm was able to extract meaningful information from rewards that were helpful to improve, and often maximize, the agent's performance.  Additionally, one important aspect of the classification results is the low percentage of false positives (Figure~\ref{fig:confusionmatrix}), showing a high specificity. On the other hand, the percentage of false negatives is generally higher.  Even though this implies that the agent misses frequently when a wrong action takes place, this is not hindering the overall performance and the agent is still learning. Although they may be scarce, accurate rewards are very useful for the RL algorithm.
